---
output:
  pdf_document:
    citation_package: natbib
    latex_engine: pdflatex
    toc: true
    toc_depth: 1
    keep_tex: true
    includes:
        in_header: vignette_head.tex
title: "GMMSensitivity: Sensitivity analysis in moment condition models"
author: "Michal Kolesár"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
bibliography: np-testing-library.bib
vignette: >
  %\VignetteIndexEntry{GMMSensitivity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, cache=FALSE}
library("knitr")
knitr::opts_knit$set(self.contained = FALSE)
knitr::opts_chunk$set(tidy = TRUE, collapse=TRUE, comment = "#>",
                      tidy.opts=list(blank=FALSE, width.cutoff=55))
```

# Introduction


The package `GMMSensitivity` implements estimators and confidence interval for
sensitivity analysis in moment condition models considered in @ak18sensitivity.
In this vignette, we demonstrate the implementation of these confidence
intervals using the demand model for automobiles from  @blp95.

The package includes the dataset `blp`, which contains estimates of the @blp95
model, as implemented by @ags17. A description of the dataset can be obtained
via the documentation in R, using `?blp`. We give additional description here:

`G`
: Matrix with 31 rows and 17 columns, estimate of derivative of the moment
  condition $G=\partial E[g(w_i,\theta)]/\partial \theta$, evaluated at initial
  estimate $\hat{\theta}_{\text{initial}}$. The initial estimate corresponds to
  a second-step GMM estimate, as computed by @ags17.

`H`
: Vector of length 17, estimate of derivative of average markup $h(\theta)$,
  evaluated at $\hat{\theta}_{\text{initial}}$.

`W`
: Weight matrix used to obtain the estimate $\hat{\theta}_{\text{initial}}$. It
  corresponds to an estimate of the inverse of the variance of the moment
  condition $\Sigma$.

`g_init`
: Average moment condition $n^{-1}\sum_{i=1}^{n}g(w_i,\theta)$, evaluated at
 estimate $\hat{\theta}_{\text{initial}}$ of $\theta$ from @blp95.

`h_init`
: Estimate of the average markup, $h(\hat{\theta}_{\text{initial}})$.

`names`
: Two lists, one for names of the moment conditions, and one for elements of theta

`ZZ`
: Gram matrix of the instruments, used to specify the set $\mathcal{C}$

`Sig`
: Variance of moment condition, estimate of $\Sigma$, given by the sample
  variance of the moment condition evaluated at $\hat{\theta}_{\text{initial}}$.

`sdZ`
: vector of standard deviations of the instruments

`perturb`
: scaling parameter to give interpretable meaning to violations of moment
 conditions. For demand-side moments, it corresponds to an estimate of
 $\delta_{d}=0.01\overline{p}/(\overline{y}/\alpha)$, and for supply-side
 moments, it corresponds to an estimate of
 $\delta_{s}=-0.01\overline{p}/\overline{m c}$, where $\overline{p}$ is average
 price, $\overline{y}$ is average income, $\overline{m c}$ is average marginal
 cost, and $\alpha$ is a parameter in the utility function. With this scaling,
 if a given demand-side instrument enters the utility function with coefficient
 $\delta_{d}$, a consumer is willing to pay 1\% of the average car price for a
 unit increase in the instrument. If a given supply-side instrument enters the
 cost function with coefficient $\delta_s$, increasing the instrument by one
 unit decreases the marginal cost by 1\% of the average car price.

`n`
: Sample size, number of model/years.


```{r, echo=FALSE}
library("GMMSensitivity")
```

# Usage

The package implements estimators, confidence intervals, and efficiency
calculations for the model (in the notation of @ak18sensitivity)
\begin{equation*}
g(\theta_{0})=c/\sqrt{n},\quad c\in \mathcal{C},\quad \mathcal{C}=\{B\gamma\colon \norm{\gamma}_{p}\leq M\}.
\end{equation*}

Suppose that we want to allow all excluded instruments in the @blp95 application
 to be potentially invalid. Fix $B$ to a scaling matrix so that if the $j$th
 supply-side instrument is invalid with $\gamma_{s j}=1$, this means that
 changing the instrument by one standard deviation changes the marginal cost by
 $\gamma_{s j}$\% of the average car price, and if the $j$ the demand-side
 instrument is invalid with $\gamma_{d j}=1$, then the consumer
 willingness to pay for one standard deviation change in the instrument
 is $\gamma_{d j}\%$ of the average 1980 car price. Let $p=2$, and $M=\sqrt{\# I}$,
 where $\#I$ is the number of invalid instruments so that $\gamma=1$ is included
 in the set (this is the same scaling as described in @ak18sensitivity in the
 paper). Then the confidence interval can be constructed as follows:

```{r}
## Construct estimate of initial sensitivity
blp$k_init <- -drop(blp$H %*% solve(crossprod(blp$G, blp$W %*% blp$G),
                                        crossprod(blp$G, blp$W)))
## list collecting initial estimates of H, G, Sigma, n, g(thetahat), initial
## sensitivity k, and initial estimate of average markup h(thetahat)
eo <- list(H=blp$H, G=blp$G, Sig=blp$Sig, n=blp$n, g_init=blp$g_init,
           k_init=blp$k_init, h_init= blp$h_init)

## Rows corresponding to invalid instruments
I <- vector(mode="logical", length=nrow(eo$G))
I[c(6:13, 20:31)] <- TRUE

## Matrix B, scaled as described in the paper
B0 <- blp$ZZ %*% diag(sqrt(blp$n)*abs(blp$perturb)/blp$sdZ)

## Value of M
M0 <- sqrt(sum(I))

## Select columns of B0 corresponding to invalid instruments
OptEstimator(eo, B0[, I], M=M0, p=2, alpha=0.05, opt.criterion="FLCI")
```

The efficiency $\kappa_{*}$ for this confidence interval can be computed using the
`EffBounds` function (which can also be used to compute efficiency of one-sided
confidence intervals):

```{r}
EffBounds(eo, B0[, I], M=M0, p=2)$twosided
```

In contrast the CI based on the initial estimate is much wider:
```{r}
OptEstimator(eo, B0[, I], M=M0, p=2, alpha=0.05, opt.criterion="Valid")
```

A specification test for whether the value $M=M0$ is too low, that is a test of
the hypothesis $H_0\colon M\leq M0$, can be conducted using the `Jtest`
function:



```{r}
## Update eo so that Sig corresponds to the initial estimate of
## Sigma, so that thetahat_initial minimizes n*g(theta)Sig^{-1} g(theta),
## and the J statistic is given by the value of this minimum.
eoJ <- eo
eoJ$Sig <- solve(blp$W)
jt <- Jtest(eoJ, B0[, I], M=M0, p=2, alpha=0.05)
```
Here `J` is the $J$-statistic, `p0` is the p-value of the usual $J$-test (that
assumes $M=0$ under the null), `pC` is the $p$-value of the test, and `Mmin` is
the smallest value of $M$ that is not rejected. Rescaling by $\sqrt{\# I}$, we
obtain the result in Table 1 in @ak18sensitivity:
```{r}
jt$Mmin/M0
```

If were only concerned about the validity of the demand-side instrument
`"demand_firm_const"` (number of cars produced by the same firm), then since
this is the sixth instrument, an analogous analysis could be conducted as:

```{r}
I <- vector(mode="logical", length=nrow(eo$G))
I[6] <- TRUE
OptEstimator(eo, B0[, I, drop=FALSE], M=1, p=2, alpha=0.05, opt.criterion="FLCI")
EffBounds(eo, B0[, I, drop=FALSE], M=1, p=2)$twosided
OptEstimator(eo, B0[, I, drop=FALSE], M=1, p=2, alpha=0.05, opt.criterion="Valid")
Jtest(eoJ, B0[, I, drop=FALSE], M=1, p=2, alpha=0.05)
```

# References
