---
title: "Sensitivty analaysis in moment condition models"
author: "Michal Kolesár"
date: "`r Sys.Date()`"

bibliography: np-testing-library.bib
output:
  pdf_document:
    toc: true
    toc_depth: 1
    includes:
        in_header: vignette_head.tex
vignette: >
  %\VignetteIndexEntry{Sensitivity analysis in GMM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, cache=FALSE}
library("knitr")
knitr::opts_knit$set(self.contained = FALSE)
knitr::opts_chunk$set(tidy = TRUE, collapse=TRUE, comment = "#>",
                      tidy.opts=list(blank=FALSE, width.cutoff=55))
```

# Introduction

The package `GMMSensitivity` implements estimators and confidence interval for
sensitivity analsysis in moment condition models considered in @ak18sensitivity.
In this vignette, we demonstrate the implementation of these confidence
intervals using the demand model for automobiles from  @blp95.

The package includes the dataset `blp`, which contains estimates of the @blp95
model, as implemented by @ags17. A description of the dataset can be obtained
using the `help` function:

```{r, echo=TRUE}
library("GMMSensitivity")
help("blp", type="text")
```

# Usage

The package implements estimators, confidence intervals, and efficiency
calculations for the model (in the notation of @ak18sensitivity)
\begin{equation*}
g(\theta_{0})=c,\quad c=B\gamma,\quad \norm{\gamma}_{p}\leq K.
\end{equation*}

For example, suppose that all excluded instruments may be invalid with
$K=\sqrt{\# I}$, where $\#I$ is the numer of invalid instruments (using the
scaling descibed in the paper). Then the confidence interval can be constructed
as follows:

```{r}
## Construct estimate of initial sensitivity
blp$k_init <- -drop(blp$H %*% solve(crossprod(blp$G, blp$W %*% blp$G),
                                        crossprod(blp$G, blp$W)))

eo <- list(H=blp$H, G=blp$G, Sig=blp$Sig, n=blp$n, g_init=blp$g_init,
           k_init=blp$k_init, h_init= blp$h_init)

## Rows corresponding to invalid instruments
I <- vector(mode="logical", length=nrow(eo$G))
I[c(6:13, 20:31)] <- TRUE

## Matrix B, scaled as described in the paper
B <- (abs(blp$perturb) * blp$OmZZ)[, I, drop=FALSE]

## Value of K
K0 <- sqrt(sum(I))

OptEstimator(eo, B, K=K0, p=2, alpha=0.05, opt.criterion="FLCI")
```
The efficiency $\kappa^*$ for this confidence interval can be computed using the
`EffBounds` function (which can also be used to compute efficiency of one-sided
confidence intervals):

```{r}
EffBounds(eo, B, K=K0, p=2)$twosided
```

For comparison, using the initial estimator yields  the confidence interval:

```{r}
OptEstimator(eo, B, K=K0, p=2, alpha=0.05, opt.criterion="Valid")
```

Finally a specification test for whether the value $K=K0$ is too low, that is a
test of the hypothesis $H_0\colon: K\leq K0$:

```{r}
Jtest(eo, B, K=K0, p=2, alpha=0.05)
```

Here `J` is the $J$-statistic, `p0` is the p-value of the usual $J$-test (that
assumes $K=0$ under the null), `pC` is the $p$-value of the test, and `Kmin` is
the smallest value of $K$ that is not rejected.

# References
